Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2015-05-26 patchy-review-responses-second-round.tex

\reply

%%%%%%
\point{}{
Regarding the simulations presented in Fig 4, I don't think they strongly support the functional form of $\migrate$. Points with very similar observed migration time have predicted migration times that differ by a factor of a 100 (the more extreme are outside the validity of the approximations, but still the I don't think this figure is strong evidence that the math is correct.)
}

\reply

%%%%%%
\point{eq.~56 is wrong}{
    as easily seen with $\Theta=0$ \ldots
    so eq.~56 has an extra 2, resulting from a wrong expansion of $\sqrt{1+x}$,
    and so eq.~57--61 are also wrong.
}

\reply
Thanks for catching that mistake.  We removed the extra factor of 2 (section~\nameref{apx:outflux}).

@@

%%%%%%
\point{time in transit}{
    One can obtain these results by a simple saddle point approximation:
    \begin{align*}
        \P\{ \text{getting to $x$ in time $t$} \} &= \P\{\text{diffusion to $x$ in time $t$}\} \P\{\text{surviving}\} \\
                                                  &= \frac{1}{(2\pi\sigma^2)^{d/2}} e^{-\frac{x^2}{2\sigma^2 t}-st}
    \end{align*}
    the exponent is minimal when
    \[
        \frac{x^2}{2\sigma^2 t^2} = s \quad \Rightarrow \quad t^* = \frac{x}{\sigma\sqrt{2s}},
    \]
    this gives your eq.~17.
    The variance is simply the inverse of the second derivative of the exponent
    \[
        \frac{1}{\var[\tau]} = \frac{x^2}{\sigma^2 (t^*)^3} = \frac{\sigma}{x(2s)^{3/2}} .
    \]
}

\reply
true, depends if you think ``saddle point approximation'' or ``laplace transform'' is a simpler idea?  this way avoids writing down more integrals; I hate to add more equations...

%%%%%%
\point{now to the thornier issues of dimension, geometry, and fluxes}{
    \begin{itemize}
        \item {} [with $\ell = \sigma/\sqrt{2s}$, $r$ the radius of the patch, $R$ the distance between patches, and $\ell \ll r \ll R$]
        \item the average density is $q(x) \sim e^{-x\sqrt{2s}/\sigma} = e^{-x/\ell}$
        \item but fluctuates far from the source as [families of size $K$, with local density $k/\ell^d$]
        \item families of size $1/s$ live for $1/s$ generations and hence contribute $1/s^2$ to the time-averaged density $q(x)$
        \item \[ q(x) = \lambda_\text{arr}(x) / s^2 \ell^d, \]
            where $\lambda_\text{arr}(x)$ is the rate at which families arrive that overlap with $x$
        \item \[ \lambda_\text{arr}(x) = q(x) s^2 \ell^d \]
            and
            \[ \Lambda_\text{arr}(x) = \lambda(x) \ell^d = q(x) s^2 \]
            would be an ``arrival density''
        \item to get at $\migrate$, arrival needs to be integrated over the relevant area. In $d=1$ this is simply a stretch of length $\ell$,
            in $d=2$ it should be $\sqrt{\ell^3 r}$,
            hence one should get
            \[ \migrate \sim \rho q(x) s^2 \ell^d \sqrt{r/\ell} (1-e^-{p_0/s}) \]
        \item result does depend on target geometry
        \item formulation in terms of fluxes might be easier
    \end{itemize}
}

@@

\reply

%%%%%%%%%%%%%%
\reviewersection

\begin{quote}
I think that the manuscript has been improved, but I don't think that all the reviewers' points have been adequately addressed. As I wrote in my previous review, I really like the authors' approach of giving simple, intuitive explanations. But writing this kind of paper actually requires a much better understanding of the model and dynamics than one needs to write a paper with the standard formal derivations, and I'm not sure that the authors are there yet.
\end{quote}

perhaps the reviewer is mistaking a different set of terminology and basic concepts for a lack of understanding?
in this process we have apparently too quickly judging physics-style arguments to be incorrect, some of the time.

%%%%%%
\point{}{
I still think that the main result (11) is dimensionally inconsistent -- it needs to be integrated over the boundary layer of the patch. In their response to my original review, the authors just referred to their response to a similar point by reviewer 1, but I don't think that this addresses my concern -- reviewer 1 was talking about integrating over a different area, the one occupied by a family, which I agree is not correct. (Note that the 1D simulations aren't going to be able to catch this issue, since you can't play with the length of the boundary.)
}

@@

\reply

%%%%%%
\point{}{
I do think that reviewer 1 raises a good point though. Looking back through the manuscript, I can't find where the authors show that the family is localized within a length scale $\sim \sigma/\sqrt{s_m} \ll A^{1/d}$. This is important, since it's needed to justify the approximation that the whole family arrives at the patch together. This seems like a basic fact about subcritical branching random walks/Brownian motions, so there's probably a formal proof of it somewhere in the literature, but I think that it's straightforward to give an intuitive explanation.
}

@@

\reply

%%%%%%
\point{}{
Speaking of the literature, what's known about hitting probabilities for subcritical branching random walks/Brownian motions?
}

@@

\reply

%%%%%%
\point{}{
I still think that the derivation of the main result is too complicated, and could be made much simpler along the lines of the suggestion in my previous review. The extra complications introduced in the current version do not actually make the result any more rigorous -- they just take a less direct route to introducing the same approximations.
}

@@

\reply

%%%%%%
\point{}{
I don't totally understand the simulation methods. The authors say that individuals first chose whether or not to reproduce with probability depending on their fitness, and then produced a Poisson-distributed number of offspring. So the selection is all in whether or not individuals reproduce, not in the number of offspring conditioned on reproduction? What exactly was the function for the probability of reproduction, and what was the mean of the offspring distribution? I'm just wondering if there's a detail here that would account for the discrepancy between the analysis and the simulations. Also, it would be helpful to say exactly what the value of sigma is in the chosen migration model.
}

@@

\reply

%%%%%%
\point{}{
I'm not sure that the description of the valid parameter range in "A review, and a test of the assumptions" is entirely correct. For the mutation results, it's fine if $s_m$ and $s_p$ have very different magnitudes, as long as the patches are sufficiently big/migration is sufficiently weak. What precisely is the condition here?
}

@@

\reply

%%%%%%
\point{}{
In the Applications, I understand that there's still a lot to learn about this system, but the authors need to at least say how far apart the actual lava flow patches are, so we have a sense of what their estimate of "tens to a few hundred kilometers" means for mutation vs migration.
}

@@

\reply

%%%%%%
\point{}{
In their response to my comments, the authors said that they did not see why the assumption $R \gg w$ was neceessary. I think that it is needed in order to talk sensibly about a single "distance to a patch" $R$: if $w \gtrsim R$, then generally some parts of the patch will be much closer than others, right?
}

@@

\reply

%%%%%%
\point{}{
I still think that including any sub-leading terms in the length of the shared haplotype is falsely precise.
}

@@

\begin{quote}
I appreciate the careful and thorough revision - and especially the exhaustive simulation work. It is an impressive and truly content-rich manuscript. I have only a few minor points:
\end{quote}

%%%%%
\point{Figure 1:}{
    point panel, $s_m = -0.02$ appears twice and $s_b$ should rather be $s_p$?
}

\reply

%%%%%
\point{Figure 4:}{
    The new figure is a bit hard to read: the grey symbols are hardly visible, square symbols appear with two meanings, the colours for the filling do not match the legend ... Maybe reduce the number of parameter combinations shown (drop some intermediate values) to focus on the "main message" and get a clearer picture? - Details can still be given in the supplement.
}

@@

\reply

%%%%%
\point{line 176:}{
    *we* provide (?)
}

\reply

%%%%%
\point{line 262:}{
    cut "depends" (?)
}

%%%%%
\point{line 263:}{
    time to reach 100 B alleles: simulations should show how long this takes $\to$ can this explain anything or is it irrelevant? (Unclear in Fig 4 since value for $s_p$ is not provided).
}

### 2015-05-27 patchy-selection-paper.tex
fixing up integration over S

To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
This is a collection of trees, each rooted at an allele living outside the patch whose parent was born inside the patch.
Next, fix some intermediate distance $r_0$ from the established patch,
and erase from the genealogy every allele that has no ancestors living further away than $r_0$ to the patch.
This fragments the original set of trees into a collection of smaller trees that relate to each other all $B$ alleles living outside of $r_0$ at any time,
and some living inside of $r_0$;
we refer to these trees as ``families''.
If $r_0$ is large enough that there are not too many families
and interactions between family members are weak,
then these families can be approximated 
by a collection of independent spatial branching processes
whose members are ignored if they enter the original patch.
(This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
In terms of these families, the equilibrium frequency above is then
\begin{align}
    \label{eqn:gestalt_q}
    \begin{split}
        \rho q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
\end{split}
\end{align}
and the quantity we wish to compute,
the effective rate at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
is
\begin{align}
    \label{eqn:gestalt_migrate}
    \begin{split}
        \migrate(x) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $x$ ) }
    \end{split}
\end{align}
The quantities on the right-hand side depend implicitly on $r_0$,
but their product does not.
All terms except the ``outflux of families'' are calculations with spatial branching processes;
and we can solve for this outflux using the established form for $q(x)$.
% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.

@@

The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
by the ratio of the second terms in the two equations,
i.e.
\begin{align} \label{eqn:ratio_gestalt}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \text{ ( probability a family establishes in patch at $x$ ) }
    }{
        \text{ ( mean occupation density of a family at $x$ ) }
    } .
\end{align}
To make this precise we need to understand what we mean by ``outflux of families'';
we do this below in section \secref{ss:hitting_occupation},
and provide a heuristic argument here.

@@

This suggests that 
\begin{align} \label{eqn:ratio_K}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \E[1-(1-p_e)^K]
    }{
        \E[K]/2s_m
    } .
\end{align}
There are now two extremes to consider.
If $\P\{K>1/p_e\}$ is small, then each family has a small chance of establishment,
and the numerator is approximately $p_e \E[K]$.
On the other hand, if $p_e$ is not too small,
then the numerator would be close to 1, as any family that arrives will likely establish.
Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
into the constant $C$,
\begin{align} \label{eqn:migrate}
  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
%     \begin{cases}
%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
%     \end{cases} .
\end{align}
When we use this below, we set $p_e = 2 s_p / \xi^2$.
Note that the approximation has the undesirable property that it is
non-monotonic in $s_m$, when we expect the rate of adaptation via
migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
which is the parameter region that we are restricted to by other assumptions.
\linelabel{rr:sm_condition}

@@

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Denote by $\gamma(x)$ the mean rate of outflux of migrant families from a point $x \in \partial B(r_0)$,
i.e.\ the time-averaged proportion of individuals near a point $x$ in $\partial B(r_0)$ 
that are the founders of new migrant families.
Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``outflux of families'',
when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
However, examining the constituent terms,
it will turn out that equation \eqref{eqn:gestalt_migrate} 
is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
Suppose that $Z$ is one such spatial branching process as above in section \secref{ss:migrant_math}, 
started at time 0 with a single individual at $x$,
and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ at location $y$
can be thought of  informally \linelabel{rr:density}
as the expected total number of offspring of a family beginning at $x$ that ever live near $y$.
It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
gives the expected total number of individuals ever living in $S$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
\end{align}
% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^\infty e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
We assume this new patch is circular, 
has relatively small area $A$ and is centered at position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
It would be possible to analyze this probability directly,
as in \citet{barton1987establishment};
but here we take a simpler route, 
approximating this by the chance that the trunk hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity
and successfully establishes in the new patch.
Write $h_A(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
\end{align}
Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
because both $q(x)$ and $\gamma(x)$ are local \emph{proportions} of $B$ alleles.

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
we show below that
\begin{align} 
    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives.
Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.

The result will not depend on $A$, 
because although $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.

Consider a single migrant family $Z$ beginning at $x$,
and write $S$ for the new patch, which has area $A$ and is centered at $y$.
Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
and $\tau_\dagger$ an independent Exponential($s_m$) time.
The mean occupation time of $Z$ spent in a region $S$ is,
since the marginal distribution of a single lineage is Brownian,
letting $u(x,S) = \int_S u(x,y) dy$,
\begin{align}
    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
\end{align}
where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
and likewise $\E^x$.
If we define
$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\tau_S}$ conditioned on $\tau_S < \tau_\dagger$,
by the strong Markov property this is equal to
\begin{align}
    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
\end{align}
where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
provided $S$ is not pathological.
If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
call this $g(A)$.

Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
(Before, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
then this is equal to
\begin{align}
    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
\end{align}
Therefore, we have that
\begin{align} \label{eqn:boils_downto}
    h(x,S) \E[K] 
    \approx \P^x[ \tau_S < \tau_\dagger ] 
    = u(x,S) / g(A) .
\end{align}

If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
\begin{align}
    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
\end{align}
Once we show that $g(A) \approx 1/(2 s_m)$, 
we will have arrived at the result, equation \eqref{eqn:ratio_K}.

>>>>

To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
This is a collection of trees, each rooted at an allele living outside the patch whose parent was born inside the patch.
Next, fix some intermediate distance $r_0$ from the established patch,
and erase from the genealogy every allele that has no ancestors living further away than $r_0$ to the patch.
This fragments the original set of trees into a collection of smaller trees that relate to each other all $B$ alleles living outside of $r_0$ at any time,
and some living inside of $r_0$;
we refer to these trees as ``families''.
If $r_0$ is large enough that there are not too many families
and interactions between family members are weak,
then these families can be approximated 
by a collection of independent spatial branching processes
whose members are ignored if they enter the original patch.
(This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
The opportunity for adaptation depends on the number $B$ alleles present at migration-selection balance:
in terms of these families, the time-averaged equilibrium number of $B$ alleles
present in the area of the new patch \emph{without} their selective advantage is,
writing $q(S) = \int_S q(y) dy$,
XXX PHRASE BETTER XXX
\begin{align}
    \label{eqn:gestalt_q}
    \begin{split}
        \rho q(S) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation in new patch per family ) } ,
\end{split}
\end{align}
and the quantity we wish to compute,
the effective rate at which \emph{families of} migrant $B$ alleles establish in the new patch
is
\begin{align}
    \label{eqn:gestalt_migrate}
    \begin{split}
        \migrate(S) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in new patch ) } .
    \end{split}
\end{align}
The quantities on the right-hand side depend implicitly on $r_0$,
but their product does not.
All terms except the ``outflux of families'' are calculations with spatial branching processes;
and we can solve for this outflux using the established form for $q(x)$.
% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.

@@

The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
suggests that we can find $\migrate(S)$ by multiplying the mean density $q(S)$
by the ratio of the second terms in the two equations,
i.e.
\begin{align} \label{eqn:ratio_gestalt}
  \migrate(S) 
    &= \rho q(S) \times 
    \frac{
        \text{ ( probability a family establishes in $S$ ) }
    }{
        \text{ ( mean occupation density of a family in $S$ ) }
    } ,
\end{align}
where again the ``occupation density'' is computed \emph{without} the selective advantage.
To make this precise we need to understand what we mean by ``outflux of families'';
we do this below in section \secref{ss:hitting_occupation},
and provide a heuristic argument here.

@@

This suggests that 
\begin{align} \label{eqn:ratio_K}
  \migrate(S) 
    &= \rho q(S) \times 
    \frac{
        \E[1-(1-p_e)^K]
    }{
        \E[K]/2s_m
    } .
\end{align}
There are now two extremes to consider.
If $\P\{K>1/p_e\}$ is small, then each family has a small chance of establishment,
and the numerator is approximately $p_e \E[K]$.
On the other hand, if $p_e$ is not too small,
then the numerator would be close to 1, as any family that arrives will likely establish.

On the other hand, XXX WHAT IS $q(S)$ XXX.

Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
into the constant $C$,
\begin{align} \label{eqn:migrate}
  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
%     \begin{cases}
%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
%     \end{cases} .
\end{align}
When we use this below, we set $p_e = 2 s_p / \xi^2$.
Note that the approximation has the undesirable property that it is
non-monotonic in $s_m$, when we expect the rate of adaptation via
migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
which is the parameter region that we are restricted to by other assumptions.
\linelabel{rr:sm_condition}

@@

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Denote by $\gamma(x)$ the mean rate of outflux of migrant families from a point $x \in \partial B(r_0)$,
i.e.\ the time-averaged density of individuals near a point $x$ in $\partial B(r_0)$ 
that are the founders of new migrant families.
Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``outflux of families'',
when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
However,
it will turn out that the integrand of equation \eqref{eqn:gestalt_migrate} 
is well-approximated by a constant multiple of the integrand of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
Suppose that $Z$ is one such spatial branching process as above in section \secref{ss:migrant_math}, 
started at time 0 with a single individual at $x$,
and write $S$ for the new patch, which has area $A$ and is centered at $y$.
The \emph{mean occupation measure} of $Z$ in the region $S$
which we denote $u(x,S)$,
% and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
% The \emph{mean occupation density} of $Z$ at location $y$
can be thought of informally \linelabel{rr:density}
as the expected total number of offspring of a family beginning at $x$ that ever live in $S$.
% It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
% i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
% gives the expected total number of individuals ever living in $S$:
% \begin{align} \label{eqn:occupation_density_defn}
%     \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
% \end{align}
% % Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
% Since reproduction and spatial motion are independent,
% $u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
% multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
% i.e.\
% \begin{align}
%     u(x,y) = \int_0^\infty e^{-s_m t} p_t(x,y) dt .
% \end{align}
Let $q(S) = \int_S q(y) dy$ denote the total equilibrium frequency in $S$.
This is decomposed in expression \eqref{eqn:gestalt_q} 
as the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    \rho q(S) &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  . 
\end{align}

Now we will decompose $u(x,S)$ under the assumption that 
the marginal distribution of the spatial motion of a single lineage is Brownian.
Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
and $\tau_\dagger$ an independent Exponential($s_m$) time.
The mean occupation time of $Z$ spent in a region $S$ is,
\begin{align}
    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
\end{align}
where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
and likewise $\E^x$.
If we define
$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\tau_S}$ conditioned on $\tau_S < \tau_\dagger$,
by the strong Markov property this is equal to
\begin{align}
    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] \\
           &= \P^x\{ \tau_S < \tau_\dagger \}  g(A)  ,
\end{align}
where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
% provided $S$ is not pathological.
and $g(A)$ is defined to be the latter expectation,
which does not depend on $x$ since $S$ is circular with area $A$.

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
It would be possible to analyze this probability directly,
as in \citet{barton1987establishment};
but here we take a simpler route, 
approximating this by the chance that the trunk hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity
and successfully establishes in the new patch.
Write $h(x,S)$ for the probability that the Brownian trunk hits the patch $S$
before the family dies out,
and $f(S)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
(This is independent of $x$ in the limit as the distance from $x$ to $S$ grows.)
As for $q(S)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(S) &= \int_{\partial B(r_0)} \gamma(x) h(x,S) f(S) dx .
\end{align}

XXX CHECK FACTORS OF $\rho$ BELOW XXX

If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
then 
\begin{align}
    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
\end{align}
Therefore, we have that
\begin{align} \label{eqn:boils_downto}
    h(x,S) \E[K] 
    \approx \P^x[ \tau_S < \tau_\dagger ] 
    = u(x,S) / g(A) .
\end{align}

% Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
% while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
% Since the expected total time spent in the patch does not depend on the distance at which the family started
% if we know that it survived to hit the patch,
% it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
% Following this intuition,
% and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
% we show below that
% \begin{align} 
%     h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
% \end{align}
% where $g(A)$ is the total expected occupation time of the patch given the family arrives.
% Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
XXX NOT RIGHT XXX
The result will not depend on $A$, 
because although $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.

By equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
\begin{align}
    \rho q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
          &= \frac{ g(A) \E[K] }{ f(A) } \migrate(S) . \label{eqn:q_migrate_relation}
\end{align}
Once we show that $g(A) \approx 1/(2 s_m)$, 
we will have arrived at the result, equation \eqref{eqn:ratio_K}.
